Parsing medial prefrontal cortex: A joint meta-analytic and graph-theoretic approach.

### 2018-08-17 Manuscript/Manuscript.Rmd
More work on manuscript

Studies of decision making have consistently associated activity in vmPFC with the subjective value assigned to decision outcomes (Bartra, McGuire, & Kable, 2010; Kable & Glimcher, 2007)

The medial prefrontal cortex (mPFC) has been suggested to subserve many of the cognitive abilities that differentiate humans from other animals (Barbas & Garcia-Cabezas, 2016). Specifically, this region has been related to decision making (Bartra, McGuire, & Kable, 2013), memory encoding (Shacter, Addis, & Buckner, 2007), and default mode deactivation (DMN) (Yeo et al., 2011), among ohters. While much has been discovered about this area in primates (Barbas & Garcia-Cabezas, 2016), the lack of direct measurements of neuronal activity makes this a tough area to record in humans, one that nowadays relies mostly on functional MRI (fMRI). Unfortunately, the mPFC is especially challenging to image, as the oxygen-related signal recorded by fMRI is contaminated by the oxygen in the sinuses (Logothetis, 2008). In addition, this area is subject to more idiosyncratic cortical folding than any other region (Zilles, Palomero- Gallagher, & Amunts, 2013), thus adding a level of complexity to the generalizable dissection of topographic functional roles within mPFC.

Previous meta analytic studies have provided important insights on the psychological phenomena attributed to mPFC subregions. For example, de la Vega et al. (2016) found subdivisions of mPFC and MCC that are significantly tied to decision making in the literature. In particular, they show that more ventral regions are more tightly associated with reward and fear. However, this pattern overlaps with their finding that vmPFC also coactivates with DMN regions, and find no topographical distinction between these phenomena. Similarly, the results from Bartra et al and Laird et al show that this region can't be subdivided into DMN vs subjective value (acikalin, chris, russ). However, meta analyses rely on peaks often derived from group-averaged data. Recent work has prescribed relevance to the analysis of single subjects in fMRI. Yeo et al (2018) propose the use of individualized topologies when identifying functional networks. Gordon et al identify important idiosincrasies in oversampled subjects. Similarly, Braga and Buckner show that finer subdivisions of the DMN are possible when looking at each subject independently. This opens up the possibility that distinctions between DMN and subjective value are possible at this finer level of resolution, but would otherwise be masked in meta analytic work.

Braga & Buckner's work is particularly informative when it comes to subdividing DMN. However, the seed-based functional connectivity analysis is sensitive to the selection of the seed. A natural next step would be to recruit all possible patterns of covariation within this network. A popular alternative is to rely on graph theoretic methods, which provide a characterization of networks that organically fits brain dynamics (provides an organic avenue to understand brain dynamics?), and thus permits the consideration of the full covariance structure of brain networks. (ADV & Tal somewhere here).

Given the shortcomings of study-specific and cross-study analyses mentioned above, there is potential value in combining complementary strengths from both approaches. Focusing on resting state data has been successful in determining a number of neural phenomena, including disentangling auditory and visual attention areas (Tobyne, Osher, Michalka, & Somers, 2017).

One great advantage of network science is the ability to characterize connectivity in a way that allows for comparisons among individuals in a meaningful way (Garcia, Ashourvan, Muldoon, Vettel, & Bassett, 2018), thus letting us address the idiosyncrasies mentioned above. In this study, we aim to subdivide mPFC into subject-specific DMN and non-DMN areas, so that we can generate more informed topographic targets for future studies of decision making. We address this aim at 3 levels of granularity: 1) across the literature (meta-analysis), selecting a number of regions of interest (ROI) that show significant overlap and specificity between DMN and decision making; 2) full-brain connectivity of actual fMRI data, using a pre-specified brain atlas; and 3) characterizing the connectivity of surface vertices (small units of function that contain oxygen-dependent neural information across time) for the regions selected in step one. I performed the last two steps at an individual subject level for 3 subjects, hoping to find some commonalities among them. In this way, I attempted to extract functional topographical information from topological network features.

>>>>

From what we talked:

1. Disentangling mPFC (unsuccessful meta-analyses)
2. Trouble imaging this area
3. Network stats in the individual as a useful method

The medial prefrontal cortex (mPFC) has been suggested to subserve many of the cognitive abilities that differentiate humans from other animals (Barbas & Garcia-Cabezas, 2016). Studies of decision making have consistently associated activity in vmPFC with the subjective value assigned to decision outcomes (Bartra, McGuire, & Kable, 2010; Kable & Glimcher, 2007). However, the distributed valuation system also includes posterior cingulate and other mPFC regions, which collectively show significant overlap with the default mode network (DMN; Fox et al., 2005; Laird et al., 2009; Yeo et al., 2011). This has prompted the question of whether these distinct phenomena are part of a common underlying brain system, or they can be dissociable.

Previous meta-analytic work has provided important insights on the psychological phenomena attributed to mPFC subregions. For example, de la Vega et al. (2016) found subdivisions of mPFC and MCC that are significantly tied to decision making across the literature. In particular, they show that more ventral mPFC regions are increasingly associated with reward and fear. However, they also found that vmPFC coactivates with DMN regions, with no topographical distinction between this network and valuation. Focused work at disentangling DMN from subjective value also failed to separate these systems (Acikalin, Gorgolewski, & Poldrack, 2017). While informative, these meta-analyses relied on peak coordinates derived from group-averaged data. It has been shown that mPFC is subject to considerable idiosyncratic cortical folding (Zilles, Palomero- Gallagher, & Amunts, 2013) and inter-subject functional variability (Mueller et al., 2013) compared to the rest of the brain, thus making group-averaging studies difficult to interpret. Aiming to remedy these shortcomings, recent work has prescribed relevance to the analysis of single subjects in fMRI. Yeo et al (2018) have effectively incorporated individualized topologies for generating distributed functional networks. Gordon et al identify important idiosincrasies in oversampled subjects that are blurred in agglomerative analyses. The trend of individualized analyses has led to finer subdivisions of the DMN (Braga & Buckner, 2018). It is thus possible that the indisitinguishable extent of DMN and valuation effects can be attributed to a lack of spatial resolution that is better understood at the individual level.

Given the shortcomings of study-specific and cross-study analyses, there is potential value in combining complementary strengths from both approaches. Focusing on resting state data has been successful in determining a number of neural phenomena, including disentangling auditory and visual attention areas (Tobyne, Osher, Michalka, & Somers, 2017). A natural next step would be to recruit all possible patterns of covariation within this network. A popular alternative is to rely on graph theoretic methods, which provide a characterization of networks that organically fits brain dynamics (provides an organic avenue to understand brain dynamics?), and thus permits the consideration of the full covariance structure of brain networks.

One great advantage of network science is the ability to characterize connectivity in a way that allows for comparisons among individuals in a meaningful way (Garcia, Ashourvan, Muldoon, Vettel, & Bassett, 2018), thus letting us address the idiosyncrasies mentioned above. In this study, we aim to subdivide mPFC into subject-specific DMN and non-DMN areas, so that we can generate more informed topographic targets for future studies of decision making. We address this aim at 3 levels of granularity: 1) across the literature (meta-analysis), selecting a number of regions of interest (ROI) that show significant overlap and specificity between DMN and decision making; and 2) characterizing the connectivity of all surface vertices for these ROIs.

Do these things truly overlap, or are we not looking close enough?

Add Tor's latest papers to the meta analysis
